Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-10-13 main.tex
biblio

We follow the approach of~\cite{Mohri03ijfcs} for the computation of distances
between words and languages
and extend it to infinite alphabets.
% with models of symbolic weighted automata and transducers.
%
The models introduced in this section generalize
weighted automata and transducers~\cite{Droste09handbook}
%over finite alphabets, see \eg~\cite{Mohri03ijfcs},
by labeling each transition with a weight function
that takes the input and output symbols as parameters.
These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
but the latter guards are restricted to the Boolean semiring.
%

@@

%
We recall that, by definition (Section~\ref{sec:prelim}),
an empty sum with $\bigoplus$ is equal to~$\zero$.
%
Intuitively,  a transition $\wei_{ij}(q, a, b, q')$ is interpreted as follows:
when reading $a$ and $b$ in the input and output words,
increment the current position in the input word if and only if $i = 1$,
and in the output word iff $j = 1$, %(otherwise, do not change it),
and change state from $q$ to $q'$.
When $a = \varepsilon$ (resp. $b = \varepsilon$), the current symbol
in the input (resp. output) is not read.
%
%In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
%the input and output symbols at current positions are always read by transitions,
%even when they do not change the reading position the head's position.
%This is an important feature in the case of an infinite alphabet in
%order to compare input and output symbols.
%which cannot be stored in the finite memory of the transducer.
%
Since~$\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and $b$.
This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for
symbolic transducers.

@@

\section{SW Visibly Pushdown Automata}
\label{section:SWVPA}\label{sec:SWVPA}
The model presented now generalizes symbolic \VPA
(\SVPA~\cite{dAntonyAlur14SVPDA},
 themselves generalizing \VPA~\cite{AlurMadhusudan09nested} to infinite alphabets)
from Boolean semirings to arbitrary semiring domains.
It associates to every nested word over an infinite alphabet a weight value
in a semiring.
Nested words can describe structures of labeled trees.
In the context of parsing, they will be useful to
represent AST % parse trees.
(see Section~\ref{sec:parsing} and Appendix~\ref{app:trees}). % parse trees. 
%\florent{see \S 5 and App.A}

@@

%\section{Nested Words and Parse Trees} \label{sec:trees}\label{app:trees}
%\input{app-trees}

>>>>

We follow the approach of~\cite{Mohri03ijfcs} for the computation of distances
between words and languages
and extend it to infinite alphabets.
% with models of symbolic weighted automata and transducers.
%
The models presented in this section generalize weighted automata and transducers~\cite{Droste09handbook}
%over finite alphabets, see \eg~\cite{Mohri03ijfcs},
by labeling each transition with a weight function
that takes the input and output symbols as parameters.
These functions generalize the guards 
of symbolic automata~\cite{Veanes12symbolic,dAntoniVeanes17CAV,dAntoni21CACM},
from Boolean domains to commutative semirings.
%
They are particular cases of the very general model of 
Weighted Symbolic Automata with Data Storage~\cite{Herrmann16dlt}.

@@

%
We recall that, by convention (Section~\ref{sec:prelim}),
an empty sum with $\bigoplus$ is equal to~$\zero$.
%
Intuitively,  a transition $\wei_{ij}(q, a, b, q')$ is interpreted as follows:
when reading $a$ and $b$ in the input and output words,
increment the current position in the input word if and only if $i = 1$,
and in the output word iff $j = 1$, %(otherwise, do not change it),
and change state from $q$ to $q'$.
When $a = \varepsilon$ (resp. $b = \varepsilon$), the current symbol
in the input (resp. output) is not read.
%
%In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
%the input and output symbols at current positions are always read by transitions,
%even when they do not change the reading position the head's position.
%This is an important feature in the case of an infinite alphabet in
%order to compare input and output symbols.
%which cannot be stored in the finite memory of the transducer.
%
Since~$\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and $b$.
This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for
symbolic transducers~\cite{Veanes12transducers}.

@@

\section{SW Visibly Pushdown Automata and Best-Search}
\label{section:SWVPA}\label{sec:SWVPA}
The model presented now generalizes symbolic \VPA
(\SVPA~\cite{dAntonyAlur14SVPDA},
 themselves generalizing \VPA~\cite{AlurMadhusudan09nested} to infinite alphabets)
from Boolean semirings to arbitrary semiring domains.
It associates to every nested word over an infinite alphabet a weight value
in a semiring.
Nested words can describe structures of labeled trees.
In the context of parsing, they will be useful to
represent AST % parse trees.
(see Section~\ref{sec:parsing} and Appendix~\ref{app:trees}). % parse trees. 
%\florent{see \S 5 and App.A}

@@

\section{Nested Words and Parse Trees} \label{sec:trees}\label{app:trees}
\input{app-trees}
